---
title: "Lab 12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing Packages
```{r}
# install.packages('ggplot2')
# install.packages('lattice')

library(ggplot2)
library(lattice)
data("mpg")

```

## Background


```{r}
##baseR
## Make a scatterplot
plot(x = mpg$displ, y = mpg$hwy, xlab = "Displacement", ylab = "Highway")
## Add loess curve
lout <- loess(hwy ~ displ, data = mpg)
ord_x <- order(lout$x)
ord_x
lines(lout$x[ord_x], lout$fitted[ord_x])
## Add two rug plots
rug(mpg$displ, side = 1)
rug(mpg$hwy, side = 2)
```

We add everything at once.
```{r}
##Lattice
xyplot(hwy ~ displ, data = mpg,
xlab = "Displacement", ylab = "Highway",
panel = function(x, y) {
panel.xyplot(x = x, y = y) ## Add scatterplot
panel.loess(x = x, y = y, span = 0.75,
family = "gaussian", degree = 2) ## Add loess curve
panel.rug(x = x, y = y) ## Add rug plot
})
```
#ggPlot2 Scatterplot
```{r}
## ggPlot2
ggplot(mpg, mapping = aes(x = displ, y = hwy)) + ## set aesthetic maps
geom_point() + ## apply aesthetic maps to
## point objects
geom_smooth(se = FALSE, method = loess) + ## apply to smooth objects
geom_rug() + ## apply to rug objects
xlab("Displacement") +
ylab("Highway")

```

##ggplot2 Foundation

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy))
#data= data frame you want to use
#mapping: definition for mapping variables onto aesthetics, almost always use aes,
#in this case mapping x and y onto x and y coordinates
```

##Comparing two quantitative variables
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point() # indicates scatterplot

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
geom_point()
```

### Annotating by Categorical Variable
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
geom_point()
#color=drv, indicates vcolor of points on scatterplot depends on value for drv
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, shape = drv)) +
geom_point()
#shape=drv indicates shape of points on scatterplot depends on value for drv

```
 Exercise: What happens when you map a categorical variable to the size or alpha aesthetics?\
 Exercise: Consider the penguins dataset that comes from the palmer penguins package:
```{r}
#install.packages("palmerpenguins")
library(palmerpenguins)
data("penguins")
```
Make a plot of bill length vs bill depth, color coding by the species of penguin. What do you notice?

```{r}
```
### Annotating by Quantitative Variable
```{r}
# Specifying quantitative  variable to determine color of point 
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = cty)) +
geom_point()
# Specifying quantitative variable to determine shading of point 
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, alpha = cty)) +
geom_point()
# Specifying quantitative variable to determine size of point 
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, size = cty)) +
geom_point()
```

• Exercise:\
What happens when we map a quantitative variable to the shape aesthetic?\

```{r}
# Shading dtermined by cty, shape by drv, color by class
ggplot(data = mpg, mapping = aes(x = displ,
y = hwy,
alpha = cty,
shape = drv,
color = class)) +
geom_point()

```

Exercise:
Using the penguins dataset, plot bill length vs bill depth, annotating by flipper length and
body mass and species. Do you notice anything?

```{r}

```
### Smoothing
```{r}
### Plotting non-parametric trend line
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_smooth(se = FALSE)

### Coloring different lines for different groups
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
geom_smooth(se = FALSE)

### Different line types for different groups
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv)) +
geom_smooth(se = FALSE)

### Default loess, plotting OLS ft by specifying method=lm
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_smooth(se = FALSE, method = lm)

### Plotting logistic curves for binary data
ggplot(data = mtcars, mapping = aes(x = hp, y = vs)) +
geom_smooth(method = glm, method.args = list(family = "binomial"), se = FALSE)
```

### Layering Geoms
```{r}
### Including both scatterplot and trendline
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE)
### Can specify aesthetic within particular geom
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))


## Should produce an error, if aesthetic not defined in ggplot, need to respecify for each geometric object
# ggplot(data = mpg) +
#   geom_point(mapping = aes(x = displ, y = hwy)) + 
#   geom_smooth()

### Specify aesthetics needed for all geom objects in ggplot (x,y), place object specific aesthetics within that object (color for geom_point)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE)
```

• Exercise:\
From the penguins dataset, try adding the OLS line to the scatterplot of bill length vs bill
depth. Color code by species, but the OLS line should not depend on species.

```{r}


  
```
### Overplotting
```{r}
## Accounting for overlapping in points by slightly nudging them around, best for small data
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
geom_jitter()

##  Accounting for overlapping by setting all points to have low transparency, best for large data
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
geom_point(alpha = 0.1)
```

• Exercise:\
Consider the diamonds dataset from ggplot2. Load it and then plot carat vs price. Account
for overplotting if possible
```{r}
data("diamonds")

```

##One Quantitative Variable
```{r}
# Histogram of hwy
ggplot(data = mpg, mapping = aes(x = hwy)) +
geom_histogram()
# Specifying number of bins for histogram
ggplot(data = mpg, mapping = aes(x = hwy)) +
geom_histogram(bins = 20)
#fregpoly- like histogram except lines specify counts
# color=drv specifies different frequency polygons for different groups, differentiated by color
ggplot(data = mpg, mapping = aes(x = hwy, color = drv)) +
geom_freqpoly(bins = 20)
## y= ..density .., plotting density instead of counts on y-axis
ggplot(data = mpg, mapping = aes(x = hwy, y = ..density.., color = drv)) +
geom_freqpoly(bins = 20)
```

Exercise:\
Create a frequency polygon of the price of a diamond, color coding by cut.
```{r}

```

### Density Plot
```{r}
###Plotting density (smoothed histogram)
ggplot(data = mpg, mapping = aes(x = hwy)) +
geom_density()

### fill=drv, different densities for different values of drv, each filkled with different color
### alpha=0.3, specifying transparency of fill
ggplot(data = mpg, mapping = aes(x = hwy, fill = drv)) +
geom_density(alpha = 0.3)
```

## One categorical variable, one quantitative variable
```{r}

```

Exercise:\
Create a boxplot for cut vs log-price. Does anything surprise you? Can you think of another
plot that could explain your surprise?
```{r}

```

## Facets
```{r}
## Creates table of scatterplots, one for each level of class
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point() +
facet_wrap( ~ class)

## Creates table of scatterplots, one for each level of class, organized in one row
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point() +
facet_grid(. ~ class)

## Creates table of scatterplots, one for each level of class, organized in one column
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point() +
facet_grid(class ~ .)

## Creates table of scatterplots, one for each combination of class and drv, each row= class, column=drv
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point() +
facet_grid(class ~ drv)
```

Exercise: \
Create a boxplot of cut vs price for each level of color.
```{r}

```

## Aesthetics for all objects

```{r}

## If you want to specify that all points have color blue, must specify color outside of aes
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
geom_point(color = "blue")

#Otherwise R assumes "blue" is categorical variable for which all observations have same value
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = "blue")) +
geom_point()

# Specifying specific properties for histogram, regardless of variable
ggplot(data = mpg, mapping = aes(x = hwy)) +
geom_histogram(fill = "white", color = "black")

```


##Themes
```{r}

## Each them specifies a different way to have background
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
geom_point() +
theme_bw()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
geom_point() +
theme_classic()

install.packages("ggthemes")
library(ggthemes)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
geom_point() +
theme_excel()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
geom_point() +
theme_fivethirtyeight()

```
• Exercise:\
Create a boxplot of cut vs price for each level of color. Make the boxplots all orange and
use the black and white theme.
```{r}

```


##Scales
```{r}
## scale_color_continuous, low= which color associates to smaller values cty, high=which color associates with larger values of cty
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = cty)) +
geom_point() +
scale_color_continuous(low = "black", high = "white")

##Scaling x and y axis by log10
ggplot(data = diamonds, mapping = aes(x = price, y = carat)) +
scale_y_log10() +
scale_x_log10() +
geom_point(alpha = 0.01)

```
## Colorblind Safe Palettes
```{r}
### scale_color_colorblind- gives points colorblind safe colors
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
geom_point() +
theme_bw() +
scale_color_colorblind()

### scale_fill_colorblind- fills density with colorblind safe colors
ggplot(data = mpg, mapping = aes(x = hwy, fill = drv)) +
geom_density(alpha = 1/3) +
theme_bw() +
scale_fill_colorblind()
```

##Labels
```{r}
## xlab= x-axis label
## ylab= y-axis label
## ggtitle= plot title
## scale_color_colorblind(name = "Car Class")= title for for key
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
geom_point() +
theme_bw() +
scale_color_colorblind(name = "Car Class") +
xlab("Displacement") +
ylab("Highway") +
ggtitle("Displacement vs Highway")

## If you dan't want to use colorblind scale, can use scale_color_discrete, scale_fill_discrete()
## scale_color_continuous, scale_fill_continuous() (depending on situation)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
geom_point() +
theme_bw() +
scale_color_discrete(name = "Car Class") +
xlab("Displacement") +
ylab("Highway") +
ggtitle("Displacement vs Highway")

## Set own colors manually
scale.colors=c(`4`="red", f="blue", r="green")
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
geom_point() +
theme_bw() +
scale_color_manual(name = "Car Drive", values=scale.colors) +
xlab("Displacement") +
ylab("Highway") +
ggtitle("Displacement vs Highway")
```
##Saving Plots
```{r}
## Storing ggplotobject in variable pl
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) +
geom_point() +
theme_bw() +
scale_color_colorblind() ->
pl

## Saving plot using ggsave
## filename= name of file ypu want to save to
## plot= ggPlot object that you want to put in pdf
## width and height of plot
## family= font for labels
ggsave(filename = "my_first_plot.pdf",
plot = pl,
width = 6,
height = 4,
family = "Times")

## pdf-opens connection to pdf file
pdf(file = "my_second_plot.pdf", width = 6, height = 4, family = "Times")
## prints saved ggPlot object  to pdf file
print(pl)
##closes connection
dev.off()
```

Exercise:\
Create a boxplot of cut vs price for each level of color. Make the boxplots all orange and
use the black and white theme. Save this plot to the output folder using ggsave().
```{r}


```